---
title: "Devo Stats Simulation"
author: "Kyle MacDonald"
date: "3/21/2017"
output: html_document
---

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(langcog)
library(lme4)
theme_set(theme_classic())
```

## Simulation functions

Function to simulate some null data for one experiment with a binary response outcome variable.

```{r}
# simulate one experiment
sim_experiment <- function(n_kids, n_trials, conditions, prob_cond_a, prob_cond_b) {
  d <- tibble()
  trial_num <- seq(1:n_trials)
  
  # simuluate kid doing task
  for(kid in 1:n_kids) {
    id <- kid
    # randomly assign to condition 
    condition <- sample(conditions, 1)
    # generate data for kid using the corresponding probability of success on each trial 
    if(condition == "a") {
      response <- replicate(n_trials, rbinom(n = 1, size = 1, prob = prob_cond_a))  
    } else {
      response <- replicate(n_trials, rbinom(n = 1, size = 1, prob = prob_cond_b))  
    }
    
    # pull kid data together in tibble
    d.tmp <- tibble(id, condition, response, trial_num)
    
    # add kid data to full tibble
    d <- bind_rows(d, d.tmp)
  }
  d
}
```

### Function to simulate the two kinds of analysis:

Treat as quantitative:
- get proportion correct for each participant
- take the mean of those proportions
- test difference between group means using t-test

Treat as binary response:
- model trial-level data using logistic regression

```{r}
sim_analysis <- function(data) {
  # standard devo-style analysis treating as means
  m_t <- data %>% 
    group_by(id, condition) %>% 
    summarise(prop_correct = mean(response)) %>% 
    t.test(prop_correct ~ condition, data = .)
  
  t <- m_t$p.value # extract p.value
  
  # logistic regression 
  logit <- data %>% 
    glm(response ~ condition, data = ., family=binomial) %>% 
    summary() %>% 
    coef %>% 
    .[2,4] # extract p.value
  
  p_df <- tibble(t, logit) %>% 
    gather(key = test, value = p_value)
}
```


## Simulation

```{r simulation}
# variables for simulation
n_obs <- 100
n_trials <- 3
n_experiments <- 100
prob_a <- 0.50
prob_b <- 0.55
cond_names <- c("a", "b")

# create empty tibble for simulation output
p_vals_df <- tibble()

for(experiment in 1:n_experiments) {
  # simulate experiment
  d <- sim_experiment(n_kids = n_obs, n_trials = n_trials, 
                      prob_cond_a = prob_a, prob_cond_b = prob_b, 
                      conditions = cond_names)
  # simulate analysis
  p_vals <- sim_analysis(d) %>% 
    mutate(exp_num = experiment)
  
  # pull together information in data frame
  p_vals_df <- bind_rows(p_vals_df, p_vals)
}
```

## Summarise and visualize simulation results

Plot the distribution of p values over the [0,1] scale

```{r}
ggplot(p_vals_df, aes(x = p_value, fill = test)) +
  geom_histogram(color = "black", position = "stack") +
  scale_fill_solarized() +
  geom_vline(xintercept = .05, size = 2, color = "darkorange", linetype = "dashed")
```

Zoom in on the decision threshold of p < .05

```{r}
p_vals_df %>% 
  filter(p_value <= 0.05) %>% 
  ggplot(., aes(x = p_value, fill = test)) +
  geom_histogram(color = "black", position = "stack") +
  scale_fill_solarized() +
  geom_vline(xintercept = .05, size = 2, color = "darkorange", linetype = "dashed")
```

Plot the total number of "significant" p values for the two analyses.

```{r}
p_vals_df %>% 
  mutate(p_sig = ifelse(p_value <= .05, T, F)) %>% 
  group_by(test) %>% 
  summarise(m = mean(p_sig)) %>% 
  ggplot(., aes(x = test, y = m, fill = test)) +
  geom_bar(stat = "identity", width = 0.5, color = "black") +
  scale_fill_solarized() 
```





